---
title: "Building and Distributing R Packages using GitHub"
subtitle: "Using GitHub for Open-Source Analytics, Reporting, and Dissemination of Research"
institute: "Center for Assessment"
author: "Damian Betebenner"
date: "April 21st, 2022 (updated: `r toOrdinal::toOrdinalDate(Sys.Date())`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: ["center", "middle", "my-title"]

---
class: inverse, center, middle

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r share-again, echo=FALSE}
xaringanExtra::use_share_again()
```

```{r style-share-again, echo=FALSE}
xaringanExtra::style_share_again(
  share_buttons = c("twitter"),
  background = "#009C49",
  foreground = "#FFFFFF"
)
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://centerforassessment.github.io/assets/favicon.png",
  width = "55px",
  height = "64px",
  position = xaringanExtra::css_position(top = "0.5em", right = "0.5em"),
  link_url = "https://www.nciea.org"
)
```

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```

```{r xaringanExtra-clipboard, echo=FALSE}
xaringanExtra::use_clipboard()
```
# R
---

# R

--

.middle[
```{r echo=FALSE}
magick::image_read_svg(path="Figs/Rlogo.svg", width=100, height=100)
```
]

--

- Open soure software environment originally developed for statistical analysis (but now useful for lots more)

--

- Has become one of the most prominent environmetns for performing and developing data analytics/statistics

--

- Applicable to many other things (this presentation and the website on which it is based was created using R)

--

- One of its greatest strengths is its extensibility

--

- Allows users to create and distribute customized R software

--

- Currently 19,002 package on CRAN and many more available only on GitHub (which is a big part of what I'm going to talk about)


---
class: inverse, center, middle

# GitHub

---

# GitHub

--

- GitHub is a code hosting platform for version control and collaboration

--

- Collaboration without coordination

--

- Often used for software development but is incredibly useful for data analytics

---

class: inverse, center, middle

# GitHub

# +

# R

# =

# Superpower

---
