---
title: "Consolidating Analysis and Dissemination of Research using GitHub"
subtitle: "Using GitHub for Open-Source Analytics, Reporting, and Dissemination of Research"
institute: "Center for Assessment"
author: "Nathan Dadey"
date: "April 21st, 2022 (updated: `r toOrdinal::toOrdinalDate(Sys.Date())`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: ["center", "middle", "my-title"]

---
class: inverse, center, middle

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r share-again, echo=FALSE}
xaringanExtra::use_share_again()
```

```{r style-share-again, echo=FALSE}
xaringanExtra::style_share_again(
  share_buttons = c("twitter"),
  background = "#009C49",
  foreground = "#FFFFFF"
)
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://centerforassessment.github.io/assets/favicon.png",
  width = "55px",
  height = "64px",
  position = xaringanExtra::css_position(top = "0.5em", right = "0.5em"),
  link_url = "https://www.nciea.org"
)
```

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```

```{r xaringanExtra-clipboard, echo=FALSE}
xaringanExtra::use_clipboard()
```


#  Motivation: Data Driven Documents 

+ In many cases, work can be shared well via one or more short vignettes or other write ups (e.g., Allie's excellent vignettes on [multiple imputation](https://centerforassessment.github.io/Internship_2021/articles/PSW_Vignette_Landing_Page.html) and [propensity score weighting](https://centerforassessment.github.io/Internship_2021/articles/MI_Vignette_Landing_Page.html))
--

+ Such write-ups can also be thought of as one form of data-driven documents (e.g., [NYT](https:/www.nytimes.com/interactive/2017/12/05/upshot/a-better-way-to-compare-public-schools.html) 
[Articles](https:/www.nytimes.com/interactive/2017/12/05/upshot/a-better-way-to-compare-public-schools.html), which typically:
  + Are well structured
  + Provide a limited n umber of comparisons
  + Meant to tell a predetermined story
--

+ This approach has been implemented in the in the prior presentations.
---

#  Motivation: Dashboards and Data Explorers 

+ In other cases, we want to encourage users to interact with the data either through dashboards (e.g., 
[CO](https://edx.cde.state.co.us/SchoolView/DataCenter/reports.jspx),
[CA](https://www.cde.ca.gov/ta/ac/cm/), 
[NYC](https://tools.nycenet.edu/), 
[MI](https://www.mischooldata.org/DistrictSchoolProfiles2/ReportCard/EducationDashboard4.aspx)) or more elaborate data exploration tools (e.g., [NAEP](https://www.nationsreportcard.gov/ndecore/landing))
--

+ These kinds of tools typically:
  + Are flexible
  + Support numerous comparisons
  + Are DIY, allowing users to find their own stories
--
  
+ Implementing this approach may involve [flexdashboard](https://rmarkdown.rstudio.com/flexdashboard/) with interactivity supported through packages like 
[DT](https://rstudio.github.io/DT/), 
[polt.ly](https://plot.ly/r/) ,
[htmlwidgets](https://www.htmlwidgets.org/) or 
Shiny (e.g., see this [simple example](https://www.lexjansen.com/pharmasug-cn/2019/DV/Pharmasug-China-2019-DV19.pdf)).
---

#  Motivation: A Third Use Case 

+ In this final case, the full complexities of multiple analysis may need to be captured (e.g., a larger set of connected write ups):
	+ Well structured across numerous write ups
	+ Contain multiple comparisons supported through multiple analyses
	+ Meant to provide a balance between structured and DIY storytelling  
--

+ One way to implement this approach is by binding up the multiple write ups from multiple analyses (captured in individual . rmds ) into a larger book via the [bookdown](https://bookdown.org/yihui/rmarkdown/books.html) package.

---
```{r pages_to_book, out.width='120%', fig.align='center', echo=FALSE, eval = TRUE}
magick::image_read("Figs/pages_to_book.png")
```

---

```{r example_workbook, out.width='120%', fig.align='center', echo=FALSE, eval = TRUE}
magick::image_read("Figs/example_workbook.png")
```

---


#  Implementation  

+ Book can be hosted via GitHub Pages by following the approach outlined in section 6.3 of Authoring Books and Technical Documents with R Markdown
[GitHub Pages](https://bookdown.org/yihui/bookdown/github.html)
[ 6.3 of Authoring Books and Technical Documents with R Markdown](https://pages.github.com/)

---

#  Considerations 

This approach may be particularly useful:
+ For analyses that are iteratively updated and shared (e.g., a workbook for state data analysis; annual psychometric analyses)
+ As a space to share final analyses, after all the messy, iterative work has been done

---

#  Supplemental Materials 



---
#  Some Orienting References 

+ Educational Data Mining (Romero & Ventura, 2010 ; Baker & Yacef, 2009; Pena-Ayala, 2014 ; Dutt, Ismail, & Herawan , 2017 )
	+ Learning Analytics (Ferguson, 2012 ; Verbert et al., 2013 )
		+ Higher Education
		+ Distance Learning
	+ Data-Driven Documents (Bostock, Ogievetsky , & Heer , 2011 )
	+ Data-Driven Decision Making (Marsh, Pane, & Hamilton, 2006 )
	+ Data Visualization
		+ Classics (Cleveland , 1994 ; Tufte , 2001 )
		+ Modern (Few, 2013 ; Nussbaumer-Knaflic , 2015 )
	+ State Report Cards (Curl & Peltzman , 2017)
