---
title: "Consolidating Analysis and Dissemination of Research using GitHub"
subtitle: "Using GitHub for Open-Source Analytics, Reporting, and Dissemination of Research"
author: "Nathan Dadey"
institute: "Center for Assessment"
date: >- 
  April 23rd, 2022<br>NCME 2022 Annual Meeting<br>Demonstration Session
  
  ```{r, github-corner, echo=FALSE, results='asis'}
    cat(cfaDocs:::makeOctoCat(
      link='https://github.com/CenterForAssessment/NCME_2022_Demonstration/blob/main/Consolidating_Analysis_and_Dissemination_of_Research_Using_GitHub.Rmd'))
  ```
output:
  cfaDocs::cfaXaringan:
    cfa_theme: null
    theme_copy: false
    css: ['**/cfa-a.css', '**/cfa-b-fonts.css']
    include: !expr list(
      footer = 'default',
      xaringanExtra = 'default',
      xExtra_use = c('share_again', 'tile_view', 'clipboard'))
---

```{r metadata, echo=FALSE}
# creating HTML metadata to accompany the slides
library(metathis)
meta() %>%
  meta_general(
    description = "The final presentation of the NCME 2022 Demonstration Session on Using GitHub for Open-Source Analytics, Reporting, and Dissemination of Research",
    subject = "Consolidating Analysis and Dissemination of Research using GitHub",
    generator = "xaringan and remark.js",
    theme_color = "#b6e202"
  ) %>% 
  meta_name("github-repo" = "centerforassessment/NCME_2022_Demonstration") %>% 
  meta_social(
    title = "Consolidating Analysis and Dissemination of Research using GitHub • NCME 2022 • Nathan Dadey",
    url = "https://centerforassessment.github.io/NCME_2022_Demonstration/articles/presentations/Demonstration_Session_Presentations.html#using-github-to-consolodate-the-analysis-and-dissemination-of-research",
    image = "https://centerforassessment.github.io/NCME_2022_Demonstration/presentations/Consolidating_Analysis_and_Dissemination_of_Research_Using_GitHub_files/TITLE_SLIDE_image.png",
    image_alt = "Title slide for Consolidating Analysis and Dissemination of Research using GitHub",
    og_type = "article",
    og_locale = "en_US",
    og_author = "Nathan Dadey",
    twitter_card_type = "summary_large_image",
    twitter_creator = "@ndadey",
    twitter_site = "@NCIEA1"
  )
```

---
class: inverse, center, middle

# Consolidating Analysis and Dissemination of Research using GitHub

---

#  Motivation: Data Driven Documents 

+ In many cases, work can be shared well via one or more **short vignettes or other write ups** (e.g., Allie's excellent vignettes on [multiple imputation](https://centerforassessment.github.io/Internship_2021/articles/PSW_Vignette_Landing_Page.html) and [propensity score weighting](https://centerforassessment.github.io/Internship_2021/articles/MI_Vignette_Landing_Page.html))
--

+ Such write-ups can also be thought of as one form of **data-driven documents** (e.g., [NYT](https:/www.nytimes.com/interactive/2017/12/05/upshot/a-better-way-to-compare-public-schools.html) 
[Articles](https:/www.nytimes.com/interactive/2017/12/05/upshot/a-better-way-to-compare-public-schools.html), which typically:
  + Are well structured
  + Provide a limited n umber of comparisons
  + Meant to tell a predetermined story
--

+ This approach has been implemented in the in the prior presentations.
---

#  Motivation: Dashboards and Data Explorers 

+ In other cases, we want to encourage users to **interact with data**, either through dashboards (e.g., 
[CO](https://edx.cde.state.co.us/SchoolView/DataCenter/reports.jspx),
[CA](https://www.cde.ca.gov/ta/ac/cm/), 
[NYC](https://tools.nycenet.edu/), 
[MI](https://www.mischooldata.org/DistrictSchoolProfiles2/ReportCard/EducationDashboard4.aspx)) or more elaborate data exploration tools (e.g., [NAEP](https://www.nationsreportcard.gov/ndecore/landing))
--

+ These kinds of tools typically:
  + Are flexible
  + Support numerous comparisons
  + Are DIY, allowing users to find their own stories
--
  
+ Implementing this approach may involve [flexdashboard](https://rmarkdown.rstudio.com/flexdashboard/) with interactivity supported through packages like 
[DT](https://rstudio.github.io/DT/), 
[polt.ly](https://plot.ly/r/) ,
[htmlwidgets](https://www.htmlwidgets.org/) or 
Shiny (e.g., see this [simple example](https://www.lexjansen.com/pharmasug-cn/2019/DV/Pharmasug-China-2019-DV19.pdf)).
---

#  Motivation: A Third Use Case 

+ In this final case, the **full complexities of multiple analysis** may need to be captured (e.g., a larger set of connected write ups):
	+ Well structured across numerous write ups
	+ Contain multiple comparisons supported through multiple analyses
	+ Meant to provide a balance between structured and DIY storytelling  
--

+ One way to implement this approach is by binding up the multiple write ups from multiple analyses (captured in individual . rmds ) into a larger book via the [bookdown](https://bookdown.org/yihui/rmarkdown/books.html) package.

---

```{r pages_to_book, fig.align='center', echo=FALSE, eval = TRUE}
magick::image_read("Figs/pages_to_book.png")
```

---
<br>
```{r example_workbook, fig.align='center', echo=FALSE, eval = TRUE}
magick::image_read("Figs/example_workbook.png")
```

---

#  Implementation  

+ Book can be hosted via GitHub Pages by following the approach outlined in section 6.3 of Authoring Books and Technical Documents with R Markdown
[GitHub Pages](https://bookdown.org/yihui/bookdown/github.html)
[ 6.3 of Authoring Books and Technical Documents with R Markdown](https://pages.github.com/)

---

#  Considerations 

This approach may be particularly useful:
+ For analyses that are **iteratively updated** and **shared** (e.g., a workbook for state data analysis; annual psychometric analyses)
--
+ As a space to share final analyses, after all the messy, iterative work has been done

---

layout: true
class: inverse, bottom

---

#  Supplemental Materials 

---

layout: false

#  Some Orienting References 

+ Educational Data Mining ([Romero & Ventura, 2010](https://ieeexplore.ieee.org/document/5524021); [Baker & Yacef, 2009; Pena-Ayala, 2014](https://jedm.educationaldatamining.org/index.php/JEDM/article/view/8); [Dutt, Ismail, & Herawan , 2017](https://ieeexplore.ieee.org/abstract/document/7820050))
	+ Learning Analytics ([Ferguson, 2012](http://oro.open.ac.uk/36374/1/IJTEL40501_Ferguson%20Jan%202013.pdf); [Verbert et al., 2013](https://journals.sagepub.com/doi/abs/10.1177/0002764213479363))
	+ Data-Driven Documents ([Bostock, Ogievetsky , & Heer , 2011](https://ieeexplore.ieee.org/abstract/document/6064996))
	+ Data-Driven Decision Making ([Marsh, Pane, & Hamilton, 2006](https://www.rand.org/pubs/occasional_papers/OP170.html))
	+ Data Visualization
		+ Classics ([Cleveland , 1994](https://www.amazon.com/Elements-Graphing-Data-William-Cleveland/dp/0963488414); [Tufte , 2001](https://www.edwardtufte.com/tufte/books_vdqi))
		+ Modern ([Few, 2013](http://www.stephen-few.com/idd.php); [Nussbaumer-Knaflic , 2015](http://www.storytellingwithdata.com/book))
	+ State Report Cards ([Curl & Peltzman , 2017](http://www.ccsso.org/sites/default/files/2017-11/CCSSO%20Reporting%20Best%20Practice%20Resource.pdf))
