---
title: "Web-based Distribution of Research Using GitHub"
subtitle: "Using GitHub for Open-Source Analytics, Reporting, and Dissemination of Research"
author: "Allie Cooperman"
institute: "University of Minnesota"
date: >- 
  April 20th, 2022<br>NCME 2022 Annual Meeting<br>Demonstration Session
  
  ```{r, github-corner, echo=FALSE, results='asis'}
    cat(cfaDocs:::makeOctoCat(
      link='https://github.com/alliecooperman'))
  ```
output:
  cfaDocs::cfaXaringan:
    include: !expr list(
      footer = 'default',
      footer_text = '<a class=\"footer\" href=\"https://centerforassessment.github.io/Internship_2021\">Research Distribution using GitHub</a>',
      xaringanExtra = 'default',
      xExtra_use = c('share_again', 'slide_tone', 'tile_view', 'clipboard'))

---

```{r metadata, echo=FALSE}
# creating HTML metadata to accompany the slides
library(metathis)
meta() %>%
  meta_general(
    description = "The second presentation of the NCME 2022 Demonstration Session on Using GitHub for Open-Source Analytics, Reporting, and Dissemination of Research",
    subject = "Web-based Distribution of Research using GitHub",
    generator = "xaringan and remark.js",
    theme_color = "#b6e202"
  ) %>% 
  meta_name("github-repo" = "centerforassessment/NCME_2022_Demonstration") %>% 
  meta_social(
    title = "Web-based Distribution of Research using GitHub • NCME 2022 • Allie Cooperman",
    url = "https://centerforassessment.github.io/NCME_2022_Demonstration/articles/presentations/Demonstration_Session_Presentations.html#web-based-distribution-of-research-using-github",
    image = "https://centerforassessment.github.io/NCME_2022_Demonstration/presentations/Web_Based_Distribution_of_Research_Using_GitHub_files/TITLE_SLIDE_image.png",
    image_alt = "Title slide for Web-based Distribution of Research using GitHub",
    og_type = "article",
    og_locale = "en_US",
    og_author = "Allie Cooperman",
    twitter_card_type = "summary_large_image",
    twitter_creator = "@acooperman",
    twitter_site = "@NCIEA1"
  )
```

class: inverse, center, middle

# Disseminating Results through <br/> GitHub Pages

---

```{css, echo = FALSE}
 h3 {
   border-bottom: none;
 }
```
# Improving Dissemination

--

### Data analysts, researchers, and practitioners are often looking for ways to clearly disseminate their findings to a broad audience.

--

### The open science push is fostering new ways to share code, data, and results that promote replication and reproducibility.

--

### [GitHub](https://github.com/) is an outstanding platform for distributing one's work, combining steps of the research project into one repository (i.e., "repo").

--

### How can we further improve the organization and presentation of a GitHub repo to more clearly disseminate one's work?

---

# GitHub Pages

--

### [GitHub Pages](https://pages.github.com/) allows users to turn their GitHub repo into a shareable, easy-to-read website.

--

### GitHub Pages is advantageous for dissemination because:

--

- Source code is easily available, facilitating reproducibility.

--

- Intensive webpage skills (e.g., HTML, JavaScript) are unnecessary, increasing accessibility to new users.

--

- GitHub is structured to foster collaboration.

---

# GitHub Pages

--

### Any public repo can be hosted as a GitHub page with a quick change of the repo settings.

--

### After cloning the repo to a local machine, users can build webpages using R and Markdown.

--

### Numerous (customizable) themes are available, including:

--
- [Supported themes from GitHub Pages](https://pages.github.com/themes/);

--
- Using the [pkgdown](https://pkgdown.r-lib.org/) R package (Wickham et al., 2022);

--
- Themes created by other GitHub users.

---

## Combining Pages with Open-Source R Packages

--

### Open-source R packages available on CRAN and GitHub allow you to more easily customize your website.

--

### The "packagePages" package from Center for Assessment was constructed by modifying "pkgdown," and builds websites through R/RStudio.

--

### The template "centerProjectTemplate" goes with "packagePages" to personalize the website design (e.g., organization colors and logos, navigation bar, headers and footers).

---

## What Can You Include in Your Website?

--
### Because GitHub Pages uses Git, R, and Markdown, numerous file types are available for your website:

--
- R Source code

--
- PDF or HTML presentations

--
- Dynamic HTML files

--

### In the context of a research project, users might push:

--
- Research findings with reproducible code

--
- Help files for R or Python functions

--
- Vignettes demonstrating different analytic methods

--
- Tutorials for running an analysis with a particular data set

---

## Graduate Student Internship Website

--
### Completed a graduate internship with the Center for Assessment in Summer 2021.

--
### Addressed methodological challenges to measuring the academic impact of COVID-19, particularly when making cross-sectional comparisons.

--
### Internship work included:

--
- Comparing the efficacy of multiple imputation methods for missing scale scores and student growth percentiles;

--
- Exploring propensity score weighting to account for school enrollment changes; and

--
- Building R Markdown scripts for automated report generation.

---

## Graduate Student Internship Website

#### Created a [GitHub-hosted website](https://centerforassessment.github.io/Internship_2021/) to showcase this work.

--

``` {r internship_homepage, out.width='75%', fig.align='center', echo=FALSE, eval = TRUE}
magick::image_read("Figs/websitehome.png")
```

---

## Graduate Student Internship Website

--

``` {r landing_page_for_propensity_score, out.width='75%', fig.align='center', echo=FALSE, eval = TRUE}
magick::image_read("Figs/pswlanding.png")
```

---

## Graduate Student Internship Website

--

### The internship website was created as a public repo, and then cloned to the local machine through an RStudio project.

--

### The "packagePages" and "centerProjectTemplate" packages allowed website customization in line with the Center for Assessment.

--

### Webpages were built through .md (Markdown) and .rmd (R Markdown) files.

--

### Source code was included wherever possible to facilitate reproducibility.

---

## Graduate Student Internship Website

--

#### Using R Markdown, we wrote a series of vignettes to present analyses throughout the internship.

--

``` {r vignette_example, out.width='70%', fig.align='center', echo=FALSE, eval = TRUE}
magick::image_read("Figs/vignetteexample.png")
```
---

## Graduate Student Internship Website

--

#### Website source code is available in the repo, but we also included Markdown files with code to recreate the vignette analyses.

--

``` {r landing_page, out.width='75%', fig.align='center', echo=FALSE, eval = TRUE}
magick::image_read("Figs/sourcecode.png")
```

---

## Graduate Student Internship Website

--

#### PDF slide decks were linked to the website for three organization-wide presentations.

--

``` {r landing_page_internship, out.width='60%', fig.align='center', echo=FALSE, eval = TRUE}
magick::image_read("Figs/presentations.png")
```

---

## Graduate Student Internship Website

--

#### Websites also allow you to include news and updates (e.g., release notes).

--

``` {r page_for_internship_updates, out.width='75%', fig.align='center', echo=FALSE, eval = TRUE}
magick::image_read("Figs/news.png")
```

---

# Resources and References

--
#### [Internship website example](https://centerforassessment.github.io/Internship_2021/index.html)

--
#### [Center for Assessment on GitHub](https://github.com/CenterForAssessment)

--
#### [Getting Started with GitHub](https://docs.github.com/en/get-started)

--
#### [GitHub Pages Documentation](https://docs.github.com/en/pages)

--
#### [pkgdown](https://pkgdown.r-lib.org/authors.html) by Wickham, Hesselberth, & Salmon (2022)

--

## Any questions? &nbsp;&nbsp;&nbsp;&nbsp;&nbsp; coope782 at umn dot edu
