---
title: "Web-based Distribution of Research Using GitHub"
subtitle: "Using GitHub for Open-Source Analytics, Reporting, and Dissemination of Research"
author: "Allie Cooperman"
institute: "University of Minnesota"
date: "April 20th, 2022"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: ["center", "middle", "my-title"]
---
class: inverse, center, middle

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r share-again, echo=FALSE}
xaringanExtra::use_share_again()
```

```{r style-share-again, echo=FALSE}
xaringanExtra::style_share_again(
  share_buttons = c("twitter"),
  background = "#009C49",
  foreground = "#FFFFFF"
)
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://centerforassessment.github.io/assets/favicon.png",
  width = "55px",
  height = "64px",
  position = xaringanExtra::css_position(top = "0.5em", right = "0.5em"),
  link_url = "https://www.nciea.org")
```

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```

```{r xaringanExtra-clipboard, echo=FALSE}
xaringanExtra::use_clipboard()
```

# Disseminating Results through GitHub Pages

---

# Improving Dissemination

--

### Data analysts, researchers, and practitioners are often looking for ways to clearly disseminate their findings to a broad audience.

--

### The open science push is fostering new ways to share code, data, and results that promote replication and reproducibility.

--

### [GitHub](https://github.com/) is an outstanding platform for distributing one's work, combining steps of the research project into one repository (i.e., "repo").

--

### How can we further improve the organization and presentation of a GitHub repo to more clearly disseminate one's work?

---

# GitHub Pages

--

### [GitHub Pages](https://pages.github.com/) allows users to turn their GitHub repo into a shareable, easy-to-read website.

--

### GitHub Pages is advantageous for dissemination because:

--

- Source code is easily available, facilitating reproducibility.

--

- Intensive webpage skills (e.g., HTML, JavaScript) are unnecessary, increasing accessibility to new users.

--

- GitHub is structured to foster collaboration.

---

# GitHub Pages

--

### Any public repo can be hosted as a GitHub page with a quick change of the repo settings.

--

### After cloning the repo to a local machine, users can build webpages using R and Markdown.

--

### Numerous (customizable) themes are available, including:

--

- [Supported themes from GitHub Pages](https://pages.github.com/themes/);

--

- Using the [pkgdown](https://pkgdown.r-lib.org/) R package (Wickham et al., 2022);

--

- Themes created by other GitHub users.

---

# Combining Pages with Open-Source R Packages

--

### Open-source R packages available on CRAN and GitHub allow you to more easily customize your website.

--

### The "packagePages" package from Center for Assessment was constructed by modifying "pkgdown," and builds websites through R/RStudio.

--

### The template "centerProjectTemplate" goes with "packagePages" to personalize the website design (e.g., organization colors and logos, navigation bar, headers and footers).

---

# What Can You Include in Your Website?

--

### Because GitHub Pages uses Git, R, and Markdown, numerous file types are available for your website:

--

- R Source code

--

- PDF or HTML presentations

--

- Dynamic HTML files

--

### In the context of a research project, users might push:

--

- Research findings with reproducible code

--

- Help files for R or Python functions

--

- Vignettes demonstrating different analytic methods

--

- Tutorials for running an analysis with a particular data set

---

# Graduate Student Internship Website

--

### Completed a graduate internship with the Center for Assessment in Summer 2021.

--

### Addressed methodological challenges to measuring the academic impact of COVID-19, particularly when making cross-sectional comparisons.

--

### Internship work included:

--

- Comparing the efficacy of multiple imputation methods for missing scale scores and student growth percentiles;

--

- Exploring propensity score weighting to account for school enrollment changes; and

--

- Building R Markdown scripts for automated report generation.

---

# Graduate Student Internship Website

### Created a [GitHub-hosted website](https://centerforassessment.github.io/Internship_2021/) to showcase this work.

--

``` {r internship_homepage, out.width='95%', fig.align='center', echo=FALSE, eval = TRUE}
magick::image_read("Figs/websitehome.png")
```

---

# Graduate Student Internship Website

--

``` {r landing_page_for_propensity_score, out.width='95%', fig.align='center', echo=FALSE, eval = TRUE}
magick::image_read("Figs/pswlanding.png")
```

---

# Graduate Student Internship Website

--

### The internship website was created as a public repo, and then cloned to the local machine through an RStudio project.

--

### The "packagePages" and "centerProjectTemplate" packages allowed website customization in line with the Center for Assessment.

--

### Webpages were built through .md (Markdown) and .rmd (R Markdown) files.

--

### Source code was included wherever possible to facilitate reproducibility.

---

# Graduate Student Internship Website

--

### Using R Markdown, we wrote a series of vignettes to present analyses throughout the internship.

--

``` {r vignette_example, out.width='95%', fig.align='center', echo=FALSE, eval = TRUE}
magick::image_read("Figs/vignetteexample.png")
```
---

# Graduate Student Internship Website

--

### Website source code is available in the repo, but we also included Markdown files with code to recreate the vignette analyses.

--

``` {r landing_page, out.width='95%', fig.align='center', echo=FALSE, eval = TRUE}
magick::image_read("Figs/sourcecode.png")
```

---

# Graduate Student Internship Website

--

### PDF slide decks were linked to the website for three organization-wide presentations.

--

``` {r landing_page_internship, out.width='75%', fig.align='center', echo=FALSE, eval = TRUE}
magick::image_read("Figs/presentations.png")
```

---

# Graduate Student Internship Website

--

### Websites also allow you to include news and updates (e.g., release notes).

--

``` {r page_for_internship_updates, out.width='95%', fig.align='center', echo=FALSE, eval = TRUE}
magick::image_read("Figs/news.png")
```

---

# Resources and References

--

### [Internship website example](https://centerforassessment.github.io/Internship_2021/index.html)

--

### [Center for Assessment on GitHub](https://github.com/CenterForAssessment)

--

### [Getting Started with GitHub](https://docs.github.com/en/get-started)

--

### [GitHub Pages Documentation](https://docs.github.com/en/pages)

--

### [pkgdown](https://pkgdown.r-lib.org/authors.html) by Wickham, Hesselberth, & Salmon (2022)

--

# Any questions?

# coope782 at umn dot edu
